PyPy's Approach to Virtual Machine Construction

### 2006-07-12 talk/dls2006/paper.tex
remove the lattice figures to get again down to 12 pages.
I'm done for today, please review my changes.

I still need to add a reference tough,
will resubmit a first time tomorrow.

However, the bytecode interpreter plays a double role, at two different
levels.  The so-called Object Spaces are \textit{domains} in the abstract
interpretation terminology.  By design, we cleanly separated these
domains from the bytecode interpreter core; the latter is only
responsible for decoding the bytecodes of an application and emulating
the corresponding stack machine.  It treats all actual application-level
objects as black boxes, and dispatches all operations on them to the
Object Space.  The Standard Object Space is a concrete domain, in which
objects are the concrete Python objects of the various built-in types:
lists, dictionaries, and so on.  By opposition, the Flow Object Space is
really an abstract domain.  It handles objects that are placeholders.
Its lattice order (actually a join-semilattice only) is shown in
figure \ref{flowlattice}.

\begin{figure*}
\centering
\includegraphics[scale=0.667]{image/flowlattice.pdf}
\caption{the lattice order of the flow object space.}
\label{flowlattice}
\end{figure*}

This order is extremely simple, because most actual analysis is delayed
to the next phase, the type inference engine.  The objects are either
\textit{Variables}, which are pure placeholders for entirely unknown values,
or \textit{Constants} with a concrete Python object as value.  The order places
Variable as the top, and keeps all \textit{Constants} unordered.  Thus if two
different constants merge during abstract interpretation, we immediately
widen them to Variable.

@@

\begin{figure*}
\centering
\includegraphics[scale=0.667]{image/lattice1.pdf}
\caption{the overall lattice.}
\label{latticeoverall}
\end{figure*}

@@

Figures \ref{latticeoverall} and \ref{latticedetail} show how these
types are ordered to form a lattice -- we generally use its structure of
join-semilattice only, but we will see in section \ref{precision}
a use case for the \textit{meet}.
Not shown on the figures are the list terms, which are simply
unordered with each other, and the Pbcs, which form a
classical finite set-of-subsets lattice.  In addition, we have left
out a number of other annotations that are irrelevant for the basic
description of the annotator and straightforward to handle:
$Dictionary$, $Tuple$, $Float$, $UnicodePoint$, $Iterator$, etc.  The
complete list is described in \cite{T}.

\begin{figure}
\centering
\includegraphics[scale=0.667]{image/lattice2.pdf}
\caption{the part about instances and nullable instances, assuming a
  simple class hierarchy with only two direct subclasses of
  \texttt{object}.}
\label{latticedetail}
\end{figure}

>>>>

However, the bytecode interpreter plays a double role, at two different
levels.  The so-called Object Spaces are \textit{domains} in the abstract
interpretation terminology.  By design, we cleanly separated these
domains from the bytecode interpreter core; the latter is only
responsible for decoding the bytecodes of an application and emulating
the corresponding stack machine.  It treats all actual application-level
objects as black boxes, and dispatches all operations on them to the
Object Space.  The Standard Object Space is a concrete domain, in which
objects are the concrete Python objects of the various built-in types:
lists, dictionaries, and so on.  By opposition, the Flow Object Space is
really an abstract domain.  It handles objects that are placeholders.

Its lattice order (actually a join-semilattice only) is extremely
simple, because most actual analysis is delayed to the next phase, the
type inference engine.  The objects are either \textit{Variables},
which are pure placeholders for entirely unknown values, or
\textit{Constants} with a concrete Python object as value.  The order
places Variable as the top, and keeps all \textit{Constants}
unordered.  Thus if two different constants merge during abstract
interpretation, we immediately widen them to Variable.

@@

@@

These types are ordered to form a lattice -- we generally use its
structure of join-semilattice only, but we will see in section
\ref{precision} a use case for the \textit{meet}. Pbcs form a
classical finite set-of-subsets lattice, class annotations are ordered
by inheritance.  We have left out a number of other annotations that
are irrelevant for the basic description of the annotator and
straightforward to handle: $Dictionary$, $Tuple$, $Float$,
$UnicodePoint$, $Iterator$, etc.  The complete list and order is
described in \cite{T}.
